### 2021-10-27 Final Doc/Chapters/Chapter2.tex

% Chapter 2

\chapter{Literature Review} % Main chapter title

\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter1}

\section{Introduction}

\section{An Issue with Instructional Education}

\section{Pedagogy and Learning Theories}
\subsection{Learner's at the Core of Learning}
\subsection{Merrill's First Principles of Instruction}
\subsection{The Influence of Motivation}

\section{An Approach Through Ludology}
\subsection{What is Ludology}
\subsection{Serious Games}
\subsection{Games as Simulation}

\section{Gamification}

\section{Existing Gamified Teaching Systems and Educational Games}

\section{Potential Effects of Game-Based Learning}

\section{Conclusion and Summary}