### 2021-11-03 Final Doc/Chapters/Chapter5.tex

% Chapter 5

\chapter{Reflection} % Main chapter title

\label{Chapter5} % For referencing the chapter elsewhere, use \ref{Chapter1}

%What did you learn while completing this project?  Refer to the decisions that you took, and evaluate them.  Analyse the strong and weak points of the product and the process that you followed.  What would you do differently?

%Did you achieve the objectives that you set for the project?  Discuss.

%How successful were you in managing the project and meeting target dates?  Discuss.

%Comments on group work.  Responsibilities and contribution of each group member.  Problems with cooperation and contributions, and dealing with these problems.

%The reflection must be structured logically.

%Conclude the reflection with a summary that is a critical evaluation of the process followed with the honours project