### 2021-11-03 Final Doc/Chapters/Chapter4.tex
Final Doc Updated

% Chapter 4

\chapter{Results} % Main chapter title

\label{Chapter4}

\section{Artefact}

\section{Article}