### 2021-11-08 Final Doc/main.tex
Updated Final Docs Abstract

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
12pt, % The default document font size, options: 10pt, 11pt, 12pt
oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
onehalfspacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
%nohyperref, % Uncomment to not load the hyperref package
headsepline, % Uncomment to get a line under the header
%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
]{MastersDoctoralThesis} % The class file specifying the document structure
\usepackage{minted}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{float}
\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international =

\usepackage{booktabs}
\usepackage{multirow}
\usepackage[final]{pdfpages}
\usepackage{mathpazo} % Use the Palatino font by default
\usepackage{changepage}
\usepackage{caption}
\usepackage[list=false]{subcaption}
\usepackage[backend=biber,style=apa,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{honsDoc.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper, % Change to letterpaper for US letter
	inner=2.5cm, % Inner margin
	outer=2cm, % Outer margin
	%bindingoffset=0.3cm, % Binding offset
	top=2cm, % Top margin
	bottom=2.5cm, % Bottom margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{The Qualities of Games for Use in Education} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Prof. Günther \textsc{Drevin}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Bachelor of Science Honours in Computer Science and Information Technology} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Joshua \textsc{Esterhuizen}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Computer Sciences} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://http://www.nwu.ac.za/}{North-West University}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{http://http://natural-sciences.nwu.ac.za/computer-sciences-and-information-systems}{School of Computer Science and Information Systems}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
%\group{\href{}{}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{http://natural-sciences.nwu.ac.za/}{Faculty of Natural and Agricultural Sciences}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\newcommand{\titleAfrikaans}{Die Eienskappe van Speletjies vir Gebruik in Onderwys}

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}
\includegraphics[scale=0.4]{Figures/nwu.png}\\[.4cm]

{\scshape\LARGE \univname\par}\vspace{0.5cm} % University name
\textsc{\Large School of Computer Science and Information Systems}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.0cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{mailto:joshua.esterhuizen27@gmail.com}{\authorname} \\ % Author name - remove the \href bracket to remove the link
\textit{30285976}
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
\href{mailto:Gunther.Drevin@nwu.ac.za}{\supname} % Supervisor name - remove the \href bracket to remove the link  
\end{flushright}
\end{minipage}\\[3cm]

\vfill

\large \textit{A thesis submitted in fulfilment of the requirements for the degree of \\ \degreename}\\[0.3cm] % University requirement text
\large \textit{for the}\\
\large \textit{ITRI671 module}\\
\vfill

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
Education is a system that had its' last major revision during the industrial age which saw instructional institutions act like factories with the goal of propelling students through the system as quick as possible. Due to this, a curriculum is often standardised and used as a means to teach all students in the same way. The issue with this is that some students may not learn effectively with the modes of education presented. One possible implementation to aid in solving this problem is the inclusion of digital and computer-based serious games into the modes of instruction.
\\\\
This study was focused on identifying possible qualities these types of games should possess to be used effectively within an educational environment. In order to reach this goal, a literature study was conducted on the fields of pedagogy, ludology and gamification with each playing a role in better understanding the means of developing these qualities. Additionally, several case studies concerning serious game artefacts were examined in which the design philosophies of the serious games were discussed and examined to identify any correlation between instructional theories. The results of this section of the project are several key qualities required as well as a brief discussion on how different topics could be approached.
\\\\
Furthermore, this project included the development of an artefact in the form of a digital computer game. These aspects of the project were conducted simultaneously with minimal correlation. The artefact was developed using the Unity Editor and C\# as the main developmental tools. The result of this is a short computer game that has a focus on simple puzzle-solving, platforming and action-based gameplay.
\\\\
\textbf{Keywords:} Game Design and Development, Gamification, Ludology, Pedagogy, Serious Games, Unity

\end{abstract}

\begin{extraAbstract}
\addchaptertocentry{\abstractnameExtra} % Add the abstract to the table of contents
Onderwys is 'n stelsel wat sy laaste groot hersiening tydens die industriële era gehad het, waarin onderriginstellings soos fabrieke opgetree het met die doel om studente so vinnig as moontlik deur die stelsel te dryf. As gevolg hiervan word 'n kurrikulum dikwels gestandaardiseer en gebruik as 'n manier om alle studente op dieselfde manier te onderrig. Die probleem hiermee is dat sommige studente dalk nie effektief leer met die onderrigmetodes wat aangebied word nie. Een moontlike implementering om te help om hierdie probleem op te los, is die insluiting van digitale en rekenaargebaseerde ernstige speletjies in die onderrigmodusse.
\\\\
Hierdie studie was gefokus op die identifisering van moontlike eienskappe wat hierdie tipe speletjies behoort te besit om effektief binne 'n opvoedkundige omgewing gebruik te word. Ten einde hierdie doel te bereik, is 'n literatuurstudie oor die velde van pedagogie, ludologie en gamification gedoen, met elkeen wat 'n rol gespeel het om die maniere om hierdie eienskappe te ontwikkel beter te verstaan. Verder is verskeie gevallestudies oor ernstige speletjie-artefakte ondersoek waarin die ontwerpfilosofieë van die ernstige speletjies bespreek en ondersoek is om enige korrelasie tussen onderrigteorieë te identifiseer. Die resultate van hierdie afdeling van die projek is verskeie sleuteleienskappe wat vereis word, asook 'n kort bespreking oor hoe verskillende onderwerpe benader kan word.
\\\\
Verder het hierdie projek die ontwikkeling van 'n artefak in die vorm van 'n digitale rekenaarspeletjie ingesluit. Hierdie aspekte van die projek is gelyktydig uitgevoer met minimale korrelasie. Die artefak is ontwikkel met behulp van die Unity Editor en C\# as die belangrikste ontwikkelingsinstrumente. Die resultaat hiervan is 'n kort rekenaarspeletjie wat 'n fokus het op eenvoudige legkaartoplossing, platforms en aksie-gebaseerde spel.
\\\\
\textbf{Keywords:} Speletjieontwerp en -ontwikkeling, Gamification, Ludologie, Pedagogie, Ernstige Spele, Unity

\end{extraAbstract}

%--------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

\listoffigures % Prints the list of figures

\listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------
%
%\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
%
%\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere\\
%\textbf{WSF} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or\\
%
%\end{abbreviations}
%
%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\include{Chapters/Chapter1}
\include{Chapters/Chapter2}
\include{Chapters/Chapter3}
\include{Chapters/Chapter4} 
\include{Chapters/Chapter5}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[heading=bibintoc,title=Reference List]

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\include{Appendices/AppendixA}
\include{Appendices/AppendixB}
\include{Appendices/AppendixC}

%----------------------------------------------------------------------------------------

\end{document}